---
title: "Data_Import"
author: "Jorid Topi"
date: "11/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(broom)
```


##Scope

This analysis further explores the carrier delay type by airline performance for this category.  


```{r}
flights_2018 <- readRDS("../app/data/flights_2018.rds")
```


# Carrier Delays

What does the distribution of carrier delays look like?

```{r}
flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  ggplot(aes(x = Carrier_Delay)) +
  geom_bar()
```

We can see that there are more carrier delays that are longer than 500 minutes. 

How is this distributed among the airlines?

```{r}
flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  ggplot(aes(x = reorder(Carrier_Name, Carrier_Delay, FUN=median), y = Carrier_Delay)) +
  geom_boxplot(aes(group = Carrier_Name)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

It doesn't seem like there are any clear patterns between the airlines. What do the means look like?

```{r}
flights_2018$Carrier_Name <- as_factor(flights_2018$Carrier_Name)

flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  group_by(Carrier_Name) %>%
  summarise(Carrier_Delay_Mean = mean(Carrier_Delay)) %>%
  mutate(Carrier_Name = fct_reorder(Carrier_Name,Carrier_Delay_Mean)) %>%
  ggplot(aes(x = Carrier_Name, y = Carrier_Delay_Mean)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

What about when adjusting for outliers? A log transformation can be used.

```{r}
flights_2018$Carrier_Name <- as_factor(flights_2018$Carrier_Name)

flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  mutate(Carrier_Delay_Log = log(Carrier_Delay)) %>%
  group_by(Carrier_Name) %>%
  summarise(Carrier_Delay_Mean_Log = mean(Carrier_Delay_Log)) %>%
  mutate(Carrier_Name = fct_reorder(Carrier_Name,Carrier_Delay_Mean_Log)) %>%
  ggplot(aes(x = Carrier_Name, y = Carrier_Delay_Mean_Log)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#Comparing Two Airlines

```{r}
carrier2 <- "Southwest Airlines Co."
carrier1 <- "SkyWest Airlines Inc."

flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  filter(Carrier_Name == carrier1 | Carrier_Name == carrier2) %>%
  ggplot(aes(x = Carrier_Name, y = Carrier_Delay)) +
  geom_boxplot(aes(group = Carrier_Name)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  filter(Carrier_Name == carrier1 | Carrier_Name == carrier2) %>%
  mutate(Carrier_Delay_Log = log(Carrier_Delay)) %>%
  t.test(Carrier_Delay_Log~Carrier_Name, data = .) %>%
  tidy() -> tout_carrier_delay_carrier

exp(tout_carrier_delay_carrier$estimate)
exp(c(low = tout_carrier_delay_carrier$conf.low, high = tout_carrier_delay_carrier$conf.high))

```
Carrier delays on SkyWest are about 2.75 times longer than carrier delays on Southwest. 





