---
title: "Data_Import"
author: "Jorid Topi"
date: "11/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(broom)
```


##Scope

This analysis explores the carrier delay type by the local hour of the scheduled departure time.  


```{r}
flights_2018 <- readRDS("../app/data/flights_2018.rds")
```

#Temporal Analysis

How do carrier delays look during the hours of the day? Do the delays get longer as the day progresses, or are airlines generally able to catch up?

```{r}
flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  group_by(Dep_Time_Planned_Local_Hour) %>%
  summarize(Carrier_Delay_Hour = mean(Carrier_Delay)) %>%
  ggplot(aes(x = Dep_Time_Planned_Local_Hour, y = Carrier_Delay_Hour)) +
  geom_bar(stat = "identity")
```

We can see that the late aircraft delay time really jumps out at the early morning hour departures. 

How does this change when log transforming to address outliers?

```{r}
flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  mutate(Carrier_Delay_log = log(Carrier_Delay)) %>%
  group_by(Dep_Time_Planned_Local_Hour) %>%
  summarize(Carrier_Delay_Hour_Log = mean(Carrier_Delay_log)) %>%
  ggplot(aes(x = Dep_Time_Planned_Local_Hour, y = Carrier_Delay_Hour_Log)) +
  geom_bar(stat = "identity")
```


How does this look for the different airlines?

```{r}
carrier1<-"JetBlue Airways"
carrier2<-"Delta Air Lines Inc."

flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  filter(Carrier_Name == carrier1 | Carrier_Name == carrier2) %>%
  group_by(Dep_Time_Planned_Local_Hour, Carrier_Name) %>%
  summarize(Carrier_Delay_Hour = mean(Carrier_Delay)) %>%
  ggplot(aes(x = Dep_Time_Planned_Local_Hour, y = Carrier_Delay_Hour, color = Carrier_Name)) +
  geom_point() + geom_smooth()
```

#Comparing Two Different Departure Times of the Day

```{r}
dep_time1 <- 5
dep_time2 <- 15

flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  filter(Dep_Time_Planned_Local_Hour == dep_time1 | Dep_Time_Planned_Local_Hour == dep_time2) %>%
  ggplot(aes(x = Dep_Time_Planned_Local_Hour, y = Carrier_Delay)) +
  geom_boxplot(aes(group = Dep_Time_Planned_Local_Hour)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

flights_2018 %>%
  filter(Carrier_Delay>0) %>%
  filter(Dep_Time_Planned_Local_Hour == dep_time1 | Dep_Time_Planned_Local_Hour == dep_time2) %>%
  mutate(Carrier_Delay_Log = log(Carrier_Delay)) %>%
  t.test(Carrier_Delay_Log~Dep_Time_Planned_Local_Hour, data = .) %>%
  tidy() -> tout_carrier_delay_hour

exp(tout_carrier_delay_hour$estimate)
exp(c(low = tout_carrier_delay_hour$conf.low, high = tout_carrier_delay_hour$conf.high))

```
Carrier delays are 2.42 times longer for depature times at the 5AM hour than for departure times at the 3PM hour. The 95% CI is 1.91 and 3.07

