---
title: "Data_Import"
author: "Jorid Topi"
date: "11/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
```


##Scope

This analysis explores the relationship between the delay types. 


```{r}
flights_2018 <- readRDS("../app/data/flights_2018.rds")
```



So how do the delay type times compare with each other?


```{r}
flights_2018 %>%
  gather("Carrier_Delay", 
             "Weather_Delay", 
             "Nas_Delay", 
             "Security_Delay", 
             "Late_Ac_Delay", 
             key = "Delay_Type", 
             value = "Delay_Type_Time") %>%
  filter(Delay_Type_Time > 0) %>%
  ggplot(aes(x = reorder(Delay_Type, Delay_Type_Time, FUN=median), y = Delay_Type_Time)) +
  geom_boxplot(aes(group = Delay_Type)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

What do the mean delay times look like?

```{r}
flights_2018 %>%
  gather("Carrier_Delay", 
             "Weather_Delay", 
             "Nas_Delay", 
             "Security_Delay", 
             "Late_Ac_Delay", 
             key = "Delay_Type", 
             value = "Delay_Type_Time") -> flights_2018_Delay_Types

flights_2018_Delay_Types$Delay_Type <- as_factor(flights_2018_Delay_Types$Delay_Type)

flights_2018_Delay_Types %>%
  filter(Delay_Type_Time>0) %>%
  group_by(Delay_Type) %>%
  summarise(Delay_Type_Mean = mean(Delay_Type_Time)) %>%
  mutate(Delay_Type = fct_reorder(Delay_Type,Delay_Type_Mean)) %>%
  ggplot(aes(x = Delay_Type, y = Delay_Type_Mean)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

It seems like on average, weather delays are the longest, with security delays the shortest. 

What does this look like when addressing outliers? A log transformation can be used.

```{r}
flights_2018_Delay_Types %>%
  filter(Delay_Type_Time>0) %>%
  mutate(Delay_Type_Time_Log = log(Delay_Type_Time)) %>%
  group_by(Delay_Type) %>%
  summarise(Delay_Type_Mean_Log = mean(Delay_Type_Time_Log)) %>%
  mutate(Delay_Type = fct_reorder(Delay_Type,Delay_Type_Mean_Log)) %>%
  ggplot(aes(x = Delay_Type, y = Delay_Type_Mean_Log)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

When adjusting for outliers, late aircraft delays are the longest on average with NAS delays the shortest.

#Comparing Delay Types

```{r}
delay_type1 <- "Nas_Delay"
delay_type2 <- "Late_Ac_Delay"

flights_2018_Delay_Types %>%
  filter(Delay_Type_Time>0) %>%
  filter(Delay_Type == delay_type1 | Delay_Type == delay_type2) %>%
  ggplot(aes(x = Delay_Type, y = Delay_Type_Time)) +
  geom_boxplot(aes(group = Delay_Type)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

flights_2018_Delay_Types %>%
  filter(Delay_Type_Time>0) %>%
  filter(Delay_Type == delay_type1 | Delay_Type == delay_type2) %>%
  mutate(Delay_Type_Time_Log = log(Delay_Type_Time)) %>%
  t.test(Delay_Type_Time_Log~Delay_Type, data = .) %>%
  tidy() -> tout_delay_type

exp(tout_delay_type$estimate)
exp(c(low = tout_delay_type$conf.low, high = tout_delay_type$conf.high))

```

It appears that Nas Delays are about 0.54 times shorter than late aircraft delays. 

