---
title: "EDA"
author: "Janice Cessna"
date: "November 30, 2019"
output: html_document
---

```{r setup and load, include=FALSE}
library(tidyverse)
library(lubridate)
library(GGally)
library(gridExtra)

flights_2018 <- readRDS("../data/flights_2018.Rds")
```



#Basic data review
Five number summary of select variables:   
```{r summary check}
summary(flights_2018[c(1, 10:18, 24:29, 31:44)])
```
Review of summaries does not immediately reveal any anomalies. Spot check of some potentially unusual variables (e.g. 420 second Air_Time and 31 minimum Distance) all checked out as seemingly legitimate.  
  
  
Total NA's in all variables:  
```{r NA tally in durations, message=FALSE, warning=FALSE}
flights_2018 %>% summarize_all(funs(sum(is.na(.)))) 
```
  
  
  
#Sub-variable creation  
This section creates a handfull of new variables for deeper EDA analysis.  
```{r creation of time sub variables}
#adds month, day of month, and weekday for each observation
flights_2018 <- flights_2018 %>%
  mutate(Date_Month = month(Date)) %>%
  mutate(Date_Day = mday(Date)) %>%
  mutate(Date_Weekday = wday(Date, label = TRUE)) 
```

```{r}
#converts air time to integer for analysis
flights_2018 <- flights_2018 %>%
  mutate(Air_Time = as.double(Air_Time) / 60)
```

```{r flights subsets}
flights_cancel <- flights_2018 %>%
  filter(Cancelled == "Yes")
flights_divert <- flights_2018 %>%
  filter(Diverted == "Yes")
```

```{r save revised RDS for use in other files}
saveRDS(flights_2018, file = "../data/flights_2018.Rds")
```

  
#Plot series: distribution of flights by different time periods
  
Date of flight: Full distribution  
```{r Date full distribution, message=FALSE, warning=FALSE}
flights_2018 %>%
  ggplot(aes(x = Date)) +
  geom_bar(binwidth = 1) 
flights_2018 %>%
  ggplot(aes(x = Date)) +
  geom_bar(binwidth = 30) 
par(mfrow = c(1,2))
```
  
```{r min-max flights per day}
#total flights per day
flights_per_day <- flights_2018 %>%
  group_by(Date) %>%
  count()
summary(flights_per_day)
flights_per_day %>%
  ggplot(aes(y = n)) +
  geom_boxplot()
#which day has the most flights
flights_per_day_max <- which(flights_per_day$n == max(flights_per_day$n))
#which day has the fewest flights
flights_per_day_min <- which(flights_per_day$n == min(flights_per_day$n))
```
The most flights on a single day occured on `r flights_per_day$Date[flights_per_day_max]` which doesn't appear to coincide with any major holiday or event. However, the fewest flights occured on `r flights_per_day$Date[flights_per_day_min]`, which happens to be U.S. Thanksgiving. These dates, along with a few other minimum date counts appear as outliers on the boxplot, but a sample of 365 days suggests their influence is probably minimal.  
```{r variable clean-up}
#clear min-max variables from global environment
rm(flights_per_day, flights_per_day_max, flights_per_day_min)
```
  
Number of flights per different time units  
```{r number flights per time units}
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Day))) +
  geom_bar() +
  xlab("Flights per day of month")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Weekday))) +
  geom_bar() +
  xlab("Flights per day of week")

flights_2018 %>%
  ggplot(aes(x = as.factor(Dep_Time_Planned_Local_Hour))) +
  geom_bar(stat = "count") +
  xlab("Flights per hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))
  
par(mfrow = c(2,2))
```
Number of flights per month not shown because that was the basis of data sampling. Thus, every month contains 10,000 flights.  
  
  
#Plot series: misc variable distributions  
  
Departure Delay: full distribution  
```{r Dep_Delay full distribution, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(flights_2018, aes(x = Dep_Delay)) +
  geom_histogram(binwidth = 10) 
```
Not surprisingly, most delays are less than 60 minutes. However, the scale of this quantity makes it hard to see what is happening in the other x-axis areas. Below is a series looking closer at these.  
  
Departure Delay: partial distributions  
```{r Dep_Delay partial distributions, echo=FALSE, message=FALSE, warning=FALSE, fig.height = 7, fig.width  = 7, fig.align  = "center"}
plot1 <- flights_2018 %>%
  filter(Dep_Delay < -20) %>%
  ggplot(aes(x = Dep_Delay)) +
  geom_histogram() +
  xlab("Dep_Delay shorter than 20 minutes")
plot2 <- flights_2018 %>%
  filter(Dep_Delay > 20) %>%
  ggplot(aes(x = Dep_Delay)) +
  geom_histogram() +
  xlab("Dep_Delay longer than 20 minutes")
plot3 <- flights_2018 %>%
  filter(Dep_Delay > 60) %>%
  ggplot(aes(x = Dep_Delay)) +
  geom_histogram() +
  xlab("Dep_Delay longer than 60 minutes")
plot4 <- flights_2018 %>%
  filter(Dep_Delay > 120) %>%
  ggplot(aes(x = Dep_Delay)) +
  geom_histogram() +
  xlab("Dep_Delay longer than 120 minutes")

grid.arrange(plot1, plot2, plot3, plot4, nrow = 2) 
rm(plot1, plot2, plot3, plot4)
```
  
Arrival Delay: full distribution  
```{r Arr_Delay full distribution, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(flights_2018, aes(x = Arr_Delay)) +
  geom_histogram(bindwidth = 10)
```
Like departure delays, most arrival delays are not surprisingly less than 60 minutes and the scale of this quantity makes it hard to see other x-axis areas. Below is a series looking closer at these.  
  
Arrival Delay: Partial distributions
```{r Arr_Delay partial distribution, echo=FALSE, message=FALSE, warning=FALSE, fig.height = 7, fig.width  = 7, fig.align  = "center"}
plot1 <- flights_2018 %>%
  filter(Arr_Delay < -20) %>%
  ggplot(aes(x = Arr_Delay)) +
  geom_histogram() +
  xlab("Arr_Delay shorter than 20 minutes")
plot2 <- flights_2018 %>%
  filter(Arr_Delay > 20) %>%
  ggplot(aes(x = Arr_Delay)) +
  geom_histogram() +
  xlab("Arr_Delay longer than 20 minutes")
plot3 <- flights_2018 %>%
  filter(Arr_Delay > 60) %>%
  ggplot(aes(x = Arr_Delay)) +
  geom_histogram() +
  xlab("Arr_Delay longer than 60 minutes")
plot4 <- flights_2018 %>%
  filter(Arr_Delay > 120) %>%
  ggplot(aes(x = Arr_Delay)) +
  geom_histogram() +
  xlab("Arr_Delay longer than 120 minutes")

grid.arrange(plot1, plot2, plot3, plot4, nrow = 2) 
rm(plot1, plot2, plot3, plot4)
```
  
Taxi-out: full distribution
```{r Taxi_Out full distribution, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(flights_2018, aes(x = Taxi_Out)) +
  geom_histogram(bindwidth = 10) 
```
  
Taxi-In: full distribution
```{r Taxi_In full distribution, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(flights_2018, aes(x = Taxi_In)) +
  geom_histogram(bindwidth = 10) 
```
  
Cancellation Code categories:  
```{r Cancellation_Code categories}
flights_2018 %>% 
  filter(!is.na(Cancellation_Code)) %>%
  ggplot(aes(x = Cancellation_Code)) +
  geom_bar()
```
  
  
#Plot series: variables by time periods

Departure delay vs explanatories  
```{r Dep_delay vs explanatories, message=FALSE, warning=FALSE}
flights_2018 %>%
  filter(Dep_Delay > 0) %>%
  ggplot(aes(x = Date, y = Dep_Delay, alpha = 0.1)) +
  geom_point() +
  theme(legend.position = "none")
flights_2018 %>%
  filter(Dep_Delay > 0) %>%
  ggplot(aes(x = as.factor(Date_Day), y = Dep_Delay, alpha = 0.1)) +
  geom_point() +
  xlab("Day of month") +
  theme(legend.position = "none")
flights_2018 %>%
  filter(Dep_Delay > 0) %>%
  ggplot(aes(x = as.factor(Date_Weekday), y = Dep_Delay, alpha = 0.1)) +
  geom_point() +
  xlab("Day of week") +
  theme(legend.position = "none")
flights_2018 %>%
  filter(Dep_Delay > 0) %>%
  ggplot(aes(x = as.factor(Dep_Time_Planned_Local_Hour), 
             y = Dep_Delay, alpha = 0.1)) +
  geom_point() +
  xlab("Delays per hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
par(mfrow = c(2,2))
```

```{r Dep_delay vs Date with Cancel code, message=FALSE, warning=FALSE}
flights_2018 %>%
  filter(Cancelled == "Yes") %>%
  ggplot(aes(x = Date, y = Dep_Delay, color = Cancellation_Code)) +
  geom_point() +
  labs(title = "Delayed flights that were then cancelled")
```
  
Arrival delay vs explanatories  
```{r Arr_delay vs explanatories, message=FALSE, warning=FALSE}
flights_2018 %>%
  filter(Arr_Delay > 0) %>%
  ggplot(aes(x = Date, y = Arr_Delay, alpha = 0.1)) +
  geom_point() +
  theme(legend.position = "none")
flights_2018 %>%
  filter(Arr_Delay > 0) %>%
  ggplot(aes(x = as.factor(Date_Day), y = Arr_Delay, alpha = 0.1)) +
  geom_point() +
  xlab("Day of month") +
  theme(legend.position = "none")
flights_2018 %>%
  filter(Arr_Delay > 0) %>%
  ggplot(aes(x = as.factor(Date_Weekday), y = Arr_Delay, alpha = 0.1)) +
  geom_point() +
  xlab("Day of week") +
  theme(legend.position = "none")
flights_2018 %>%
  filter(Arr_Delay > 0) %>%
  ggplot(aes(x = as.factor(Arr_Time_Planned_Local_Hour), 
             y = Arr_Delay, alpha = 0.1)) +
  geom_point() +
  xlab("Delays per hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
par(mfrow = c(2,2))
```

Air time per different time units  
```{r air time per time units, message=FALSE, warning=FALSE}
flights_2018 %>%
  ggplot(aes(x = Date, y = Air_Time, alpha = 0.1)) +
  geom_point() +
  theme(legend.position = "none")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Day), y = Air_Time, alpha = 0.1)) +
  geom_point() +
  xlab("Day of month") +
  theme(legend.position = "none")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Weekday), y = Air_Time, alpha = 0.1)) +
  geom_point() +
  xlab("Day of week") +
  theme(legend.position = "none")

flights_2018 %>%
  ggplot(aes(x = as.factor(Dep_Time_Planned_Local_Hour),
             y = Air_Time, alpha = 0.1)) +
  geom_point() +
  xlab("Hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
  
par(mfrow = c(2,2))
```
  
Distance per different time units  
```{r distance per time units, message=FALSE, warning=FALSE}
flights_2018 %>%
  ggplot(aes(x = Date, y = Distance, alpha = 0.1)) +
  geom_point() +
  theme(legend.position = "none")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Day), y = Distance, alpha = 0.1)) +
  geom_point() +
  xlab("Day of month") +
  theme(legend.position = "none")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Weekday), y = Distance, alpha = 0.1)) +
  geom_point() +
  xlab("Day of week") +
  theme(legend.position = "none")

flights_2018 %>%
  ggplot(aes(x = as.factor(Dep_Time_Planned_Local_Hour),
             y = Distance, alpha = 0.1)) +
  geom_point() +
  xlab("Hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
  
par(mfrow = c(2,2))
```
  
Taxi-out per different time units  
```{r taxi-out per time units, message=FALSE, warning=FALSE}
flights_2018 %>%
  ggplot(aes(x = Date, y = Taxi_Out, alpha = 0.1)) +
  geom_point() +
  theme(legend.position = "none")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Day), y = Taxi_Out, alpha = 0.1)) +
  geom_point() +
  xlab("Day of month") +
  theme(legend.position = "none")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Weekday), y = Taxi_Out, alpha = 0.1)) +
  geom_point() +
  xlab("Day of week") +
  theme(legend.position = "none")
  
flights_2018 %>%
  ggplot(aes(x = as.factor(Dep_Time_Planned_Local_Hour),
             y = Taxi_Out, alpha = 0.1)) +
  geom_point() +
  xlab("Hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
  
par(mfrow = c(2,2))
```
  
Taxi-in per different time units  
```{r taxi-in per time units, message=FALSE, warning=FALSE}
flights_2018 %>%
  ggplot(aes(x = Date, y = Taxi_In, alpha = 0.1)) +
  geom_point() +
  theme(legend.position = "none")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Day), y = Taxi_In, alpha = 0.1)) +
  geom_point() +
  xlab("Day of month") +
  theme(legend.position = "none")
flights_2018 %>%
  ggplot(aes(x = as.factor(Date_Weekday), y = Taxi_In, alpha = 0.1)) +
  geom_point() +
  xlab("Day of week") +
  theme(legend.position = "none")

flights_2018 %>%
  ggplot(aes(x = as.factor(Dep_Time_Planned_Local_Hour),
             y = Taxi_In, alpha = 0.1)) +
  geom_point() +
  xlab("Hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
  
par(mfrow = c(2,2))
```
 
Cancelled: distributions
```{r Cancelled distributions, message=FALSE, warning=FALSE}
flights_2018 %>%
  filter(Cancelled == "Yes") %>%
  ggplot(aes(x = Date)) +
  geom_bar(aes(fill = Cancellation_Code))
flights_2018 %>%
  filter(Cancelled == "Yes") %>%
  ggplot(aes(x = as.factor(Date_Day))) +
  geom_bar() +
  xlab("Cancellations per day of month")
flights_2018 %>%
  filter(Cancelled == "Yes") %>%
  ggplot(aes(x = as.factor(Date_Weekday))) +
  geom_bar() +
  xlab("Cancellations per day of week") 
flights_2018 %>%
  filter(Cancelled == "Yes") %>%
  ggplot(aes(x = as.factor(Arr_Time_Planned_Local_Hour))) +
  geom_bar() +
  xlab("Cancellations per hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
par(mfrow = c(2,2))
```
  
Diversion: distributions 
```{r Diversion distributions, message=FALSE, warning=FALSE}
flights_2018 %>%
  filter(Diverted == "Yes") %>%
  ggplot(aes(x = Date)) +
  geom_bar()
flights_2018 %>%
  filter(Diverted == "Yes") %>%
  ggplot(aes(x = as.factor(Date_Day))) +
  geom_bar() +
  xlab("Diversions per day of month")
flights_2018 %>%
  filter(Diverted == "Yes") %>%
  ggplot(aes(x = as.factor(Date_Weekday))) +
  geom_bar() +
  xlab("Diversions per day of week") 
flights_2018 %>%
  filter(Diverted == "Yes") %>%
  ggplot(aes(x = as.factor(Arr_Time_Planned_Local_Hour))) +
  geom_bar() +
  xlab("Diversions per hour of day") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
par(mfrow = c(2,2))
```
 
