---
title: "EDA_stat"
author: "Janice Cessna"
date: "November 30, 2019"
output: html_document
---

```{r}
library(tidyverse)
library(broom)
library(pls)

flights_2018 <- readRDS("../data/flights_2018.Rds")
```


```{r select data for analysis}
#flights_2018 <- flights_2018 %>%
#  filter(Cancelled == "No") %>%
#  filter(Diverted == "No") 
```



##How does Airport performance look at different times of the day? 
```{r}

```


```{r}
mean_delay_overall <- mean(flights_2018$Dep_Delay)
```
The overall mean departure delay for all 2018 flights was `r mean_delay_overall`. When we compare the mean delay for each hour of day segment (H0: mu = `r mean_delay_overall`), we find that... 
```{r}
flights_nested <- flights_2018 %>%
  filter(Cancelled == "No") %>%
  filter(Diverted == "No") %>%  
  select(Dep_Delay, Dep_Time_Planned_Local_Hour) %>%
  mutate(Diff_from_mean = Dep_Delay - mean_delay_overall) %>%
  arrange(desc(Dep_Time_Planned_Local_Hour)) %>%
  group_by(Dep_Time_Planned_Local_Hour) %>%
  nest()

#code below here does not work yet
flights_nested %>%
  ggplot(aes(x = data[Dep_Time_Planned_Local_Hour], y = data[Diff_from_mean])) +
  geom_boxplot()


flights_nested <- flights_nested %>%
  mutate(tout = map(data, ~t.test(Diff_from_mean, data = .))) 
```



##Is this pattern consistent throughout the year? 
```{r}

```


















#Multiple linear regression 
```{r multiple linear regression}
mls_fit <- lm(Dep_Delay ~ ., data = flights_2018_EDAstat2)
mls_tidy <- tidy(mls_fit)
mls_augment <- augment(mls_fit)
mls_tidy
mls_augment
```
```{r}
mls_augment %>%
  ggplot(aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = 2)
```



#Principle components regression
```{r principle components regression}
#set.seed(2)
flights_2018_pcr <- flights_2018_EDAstat2 %>%
  select(-c(1:6))
pcr_fit <- pcr(Dep_Delay ~., data = flights_2018_pcr, scale = T, validation = "CV")
summary(pcr_fit)
```

```{r elbow plots}
validationplot(pcr_fit, val.type="MSEP", legendpos="topright") # Plots the MSE
validationplot(pcr_fit, val.type="RMSEP", legendpos="bottomright") # Plots the SquareRoot(MSE) 
```

```{r loadings summary}
str(pcr_fit)
pcr_fit$loadings
```

```{r PCR with 12 components}
pcr_fit2 <- pcr(Dep_Delay ~ ., data = flights_2018_pcr, scale = T, ncomp = 12) 
summary(pcr_fit2)
```



#Partial least squares regression
```{r}
library(pls)

pls_fit <- plsr(Dep_Delay ~ ., data = flights_2018_pcr, scale = T, validation = "CV")
summary(pls_fit)
```

```{r}
validationplot(pls_fit, val.type = "MSEP", legendpos = "topright") # w/MSE
validationplot(pls_fit, val.type = "RMSEP", legendpos = "topright") # w/RMSE
```

```{r}
pls_fit$loadings
```

```{r}
pls_fit2 <-plsr(Dep_Delay ~ ., data = flights_2018_pcr, scale = T, ncomp = 2) 
summary(pls_fit2)
```



#PCR and PLS displayed together
```{r}
cbind("PCR RMSE"=sqrt(mean((pcr.pred-Hitters[pred.test,]$Salary)^2)), "PLS RMSE"= sqrt(mean((pls.pred-Hitters[pred.test,]$Salary)^2)))
```

