---
title: "Data_Import"
author: "Jorid Topi"
date: "11/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(broom)
```


##Scope

This analysis further explores the late aircraft type delays. How does this compare among the different airlines?


```{r}
flights_2018 <- readRDS("../app/data/flights_2018.rds")
```


#Late Aircraft Delays

What does the distribution of late aircraft delays look like?

```{r}
flights_2018 %>%
  filter(Late_Ac_Delay>0 & Late_Ac_Delay<400) %>%
  ggplot(aes(x = Late_Ac_Delay)) +
  geom_bar()
```

How is this distributed among the airlines?

```{r}
flights_2018 %>%
  filter(Late_Ac_Delay>0) %>%
  ggplot(aes(x = reorder(Carrier_Name, Late_Ac_Delay, FUN = median), y = Late_Ac_Delay)) +
  geom_boxplot(aes(group = Carrier_Name)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

It doesn't seem like there are any clear patterns between the airlines. What do the means look like?

```{r}
flights_2018$Carrier_Name <- as_factor(flights_2018$Carrier_Name)

flights_2018 %>%
  filter(Late_Ac_Delay>0) %>%
  group_by(Carrier_Name) %>%
  summarise(Late_Ac_Mean = mean(Late_Ac_Delay)) %>%
  mutate(Carrier_Name = fct_reorder(Carrier_Name,Late_Ac_Mean)) %>%
  ggplot(aes(x = Carrier_Name, y = Late_Ac_Mean)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

What about when adjusting for outliers? A log transformation can be used.

```{r}
flights_2018$Carrier_Name <- as_factor(flights_2018$Carrier_Name)

flights_2018 %>%
  filter(Late_Ac_Delay>0) %>%
  mutate(Late_Ac_Delay_Log = log(Late_Ac_Delay)) %>%
  group_by(Carrier_Name) %>%
  summarise(Late_Ac_Mean_Log = mean(Late_Ac_Delay_Log)) %>%
  mutate(Carrier_Name = fct_reorder(Carrier_Name,Late_Ac_Mean_Log)) %>%
  ggplot(aes(x = Carrier_Name, y = Late_Ac_Mean_Log)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


#Comparing two airlines

```{r}
carrier1 <- "Hawaiian Airlines Inc."
carrier2 <- "Mesa Airlines Inc."

flights_2018 %>%
  filter(Late_Ac_Delay>0) %>%
  filter(Carrier_Name == carrier1 | Carrier_Name == carrier2) %>%
  ggplot(aes(x = Carrier_Name, y = Late_Ac_Delay)) +
  geom_boxplot(aes(group = Carrier_Name)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

flights_2018 %>%
  filter(Late_Ac_Delay>0) %>%
  filter(Carrier_Name == carrier1 | Carrier_Name == carrier2) %>%
  mutate(Late_Ac_Delay_Log = log(Late_Ac_Delay)) %>%
  t.test(Late_Ac_Delay_Log~Carrier_Name, data = .) %>%
  tidy() -> tout_late_ac_delay_carrier

exp(tout_late_ac_delay_carrier$estimate)
exp(c(low = tout_late_ac_delay_carrier$conf.low, high = tout_late_ac_delay_carrier$conf.high))
```

Late aircraft delays seem to be 2.23 times longer for Mesa Airlines than for Hawaiian. The 95% confidence interval is 1.75, 2.84



