---
title: "Regression: Security_Delay on Total Ops"
author: "Janice Cessna"
date: "December 5, 2019"
output: html_document
urlcolor: "blue"
params:
  solutions: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = params$solutions, 
                      eval       = params$solutions)
```

    
```{r, echo = TRUE}
library(tidyverse)
library(broom)
```
```{r}
flights_2018 <- readRDS("../app/data/flights_2018.rds")
```
    

#Data setup

Establishing data frame with reduced set of initial variables and creation of additional variables for regression models below. 
```{r}
flights_2018_SD <- flights_2018 %>%
  select(Origin,Dep_Time_Planned_Local_Hour, Dep_Delay, 
         Taxi_Out, Taxi_In, Carrier_Delay, Weather_Delay, Nas_Delay, 
         Security_Delay, Late_Ac_Delay, Tot_Ops.x) %>%
  filter(Security_Delay > 0) %>%  #changed per y variable
  group_by(Origin) %>% 
  mutate(Mean_Dep_Delay = mean(Dep_Delay)) %>%
  mutate(Mean_Taxi_Out = mean(Taxi_Out)) %>%
  mutate(Mean_Taxi_In = mean(Taxi_In)) %>%
  mutate(Mean_Carrier_Delay = mean(Carrier_Delay)) %>%
  mutate(Mean_Weather_Delay = mean(Weather_Delay)) %>%
  mutate(Mean_Nas_Delay = mean(Nas_Delay)) %>%
  mutate(Mean_Security_Delay = mean(Security_Delay)) %>%
  mutate(Mean_Late_Ac_Delay = mean(Late_Ac_Delay)) %>%
  mutate(Duplicate_Origin = duplicated(Origin)) %>%
  filter(Duplicate_Origin == "FALSE")
```


# Regression Model

Plotting mean security delay versus total operations by airport. 
```{r}
flights_2018_SD %>%
  ggplot(aes(x = Tot_Ops.x, y = Mean_Security_Delay)) +
  geom_point(na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE, na.rm = TRUE) +
  xlab("Total Airport Operations") +
  ylab("Mean Security Delay") +
  labs(title = "Delay trend by operations") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))
```

Linear regression of mean security delay on total operations by airport suggests a slightly decreasing security delay as total operations increases. 
```{r}
lm_Security_delay <- lm(Mean_Security_Delay ~ Tot_Ops.x, data = flights_2018_SD)
lm_Security_delay_tidy <- tidy(lm_Security_delay)
lm_Security_delay_tidy
lm_Security_delay_augment <- augment(lm_Security_delay)
lm_Security_delay_augment
```

A plot of the regression residuals shows a linear relationship, although there seems to be a skew for all x.  
```{r}
lm_Security_delay_augment %>%
  ggplot(aes(x = .fitted, y = .std.resid)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```

After logging y, the plot of mean security delay on total operations now looks like:   
```{r}
flights_2018_SD %>%
  ggplot(aes(x = Tot_Ops.x, y = log2(Mean_Security_Delay))) +
  geom_point(na.rm = TRUE) +
  geom_smooth(method = "lm", se = FALSE, na.rm = TRUE) +
  xlab("Total Airport Operations") +
  ylab("Mean Security Delay") +
  labs(title = "Delay trend by operations") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0))
```

The revised linear regression model is now: 
```{r}
lm_Security_delay <- lm(log2(Mean_Security_Delay) ~ Tot_Ops.x, data = flights_2018_SD)
lm_Security_delay_tidy <- tidy(lm_Security_delay)
lm_Security_delay_tidy
lm_Security_delay_augment <- augment(lm_Security_delay)
```

And the revised plot of residuals shows variance is nicely distributed. 
```{r}
lm_Security_delay_augment %>%
  ggplot(aes(x = .fitted, y = .std.resid)) +
  geom_point() +
  geom_smooth(method = lm, se = FALSE)
```


