---
title: "Regression: Dep_Delay on Total Ops"
author: "Janice Cessna"
date: "December 5, 2019"
output: html_document
urlcolor: "blue"
params:
  solutions: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = params$solutions, 
                      eval       = params$solutions)
```

    
```{r, echo = TRUE}
library(tidyverse)
library(broom)
```
```{r}
flights_2018 <- readRDS("../app/data/flights_2018.rds")
```
    

#Data setup

Establishing data frame with reduced set of initial variables and creation of additional variables for regression models below. 
```{r}
flights_temp <- flights_2018 %>%
  select(Origin,Origin_City, Origin_State, Region,
         Dep_Time_Planned_Local_Hour, Dep_Delay, 
         Taxi_Out, Taxi_In, Carrier_Delay, Weather_Delay, Nas_Delay, 
         Security_Delay, Late_Ac_Delay, Tot_Ops.x) %>%
  filter(Dep_Delay > 0) %>% 
  group_by(Origin) %>% 
  mutate(Mean_Dep_Delay = mean(Dep_Delay)) %>%
  mutate(Mean_Taxi_Out = mean(Taxi_Out)) %>%
  mutate(Mean_Taxi_In = mean(Taxi_In)) %>%
  mutate(Mean_Carrier_Delay = mean(Carrier_Delay)) %>%
  mutate(Mean_Weather_Delay = mean(Weather_Delay)) %>%
  mutate(Mean_Nas_Delay = mean(Nas_Delay)) %>%
  mutate(Mean_Security_Delay = mean(Security_Delay)) %>%
  mutate(Mean_Late_Ac_Delay = mean(Late_Ac_Delay)) %>%
  mutate(Duplicate_Origin = duplicated(Origin)) %>%
  filter(Duplicate_Origin == "FALSE")
```


# Regression Model

Plotting mean departure delay versus total operations by airport and grouping by FAA region. 
```{r}
flights_temp %>%
  ggplot(aes(x = Tot_Ops.x, y = Mean_Dep_Delay, group = Region, color = Region)) +
  geom_point(alpha = 0.5) +
  xlab("Total Airport Operations") +
  ylab("Mean Departure Delay") +
  labs(title = "Delay trend by operations") 
```

Linear regression model output for mean departure delay on total operations by region: 
```{r}
flights_nested <- flights_temp %>%
  group_by(Region) %>%
  nest() 

flights_nested <- flights_nested %>%
  mutate(lmout = map(data, ~lm(Mean_Dep_Delay ~ Tot_Ops.x, data = .))) %>%
  mutate(tidyout = map(lmout, ~tidy(., conf.int = TRUE)))

flights_lmtidy <- unnest(flights_nested, tidyout) %>%
  filter(term == "Tot_Ops.x") %>%
  select(c("Region", "estimate", "p.value", "conf.low", "conf.high"))
flights_lmtidy
```

```{r eval=FALSE, include=FALSE}
#ignore this bit -- just experimenting
flights_lmtidy %>%
  filter (term == "Tot_Ops.x") %>%
  mutate(region_order = fct_reorder(Region, estimate)) %>%
  ggplot(aes(x = region_order, y = estimate)) +
  geom_point() +
  geom_hline(yintercept = c(-0.01,0,0.01), linetype = 2) +
  geom_segment(aes(x = region_order, xend = region_order, y = conf.low, yend = conf.high), 
               color = "black", alpha = 0.3) +
  ylab("Rate of  change") +
  xlab("Region") +
  labs(title = "Average estimated departure delay per region with confindence intervals") +
  theme(axis.text.x = element_blank(),axis.ticks.x = element_blank())
```

```{r}
flights_lmtidy %>%
  ggplot(aes(x = fct_reorder(Region, estimate), y = estimate)) +
  geom_point(aes(color = cut(p.value, c(-Inf, 0.05, Inf)))) +
  geom_hline(yintercept = 0, color = "red", linetype = 2) +
  geom_segment(aes(x = Region, xend = Region, y = conf.low, yend = conf.high), 
               color = "black", alpha = 0.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0)) +
  theme(legend.position = "none")
```
The estimated mean departure delays are not much different than zero and their confidence intervals all include zero. 




