---
title: "Data_Import"
author: "Jorid Topi"
date: "11/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(broom)
```


##Scope

Late aircraft and carrier delays are the responsibility of the airline. As such, a combined analysis may be beneficial for exploring airline performance.


```{r}
flights_2018 <- readRDS("../app/data/flights_2018.rds")
```

#Late Aircraft and Carrier Delays Combined

Since both these types of delays are due to the airline responsibility, a combined analysis may be beneficial.

What does the distribution of carrier and late aircraft delays look like?

```{r}
flights_2018 %>%
  mutate(Total_Delay = Carrier_Delay + Late_Ac_Delay) %>%
  filter(Total_Delay>0) %>%
  ggplot(aes(x = Total_Delay)) +
  geom_bar()
```

How is this distributed among the airlines?

```{r}
flights_2018 %>%
  mutate(Total_Delay = Carrier_Delay + Late_Ac_Delay) %>%
  filter(Total_Delay>0) %>%
  ggplot(aes(x = reorder(Carrier_Name, Total_Delay, FUN = median), y = Total_Delay)) +
  geom_boxplot(aes(group = Carrier_Name)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

It doesn't seem like there are any clear patterns between the airlines. What do the means look like?

```{r}
flights_2018$Carrier_Name <- as_factor(flights_2018$Carrier_Name)

flights_2018 %>%
  mutate(Total_Delay = Carrier_Delay + Late_Ac_Delay) %>%
  filter(Total_Delay>0) %>%
  group_by(Carrier_Name) %>%
  summarise(Total_Delay_Mean = mean(Total_Delay)) %>%
  mutate(Carrier_Name = fct_reorder(Carrier_Name,Total_Delay_Mean)) %>%
  ggplot(aes(x = Carrier_Name, y = Total_Delay_Mean)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

What about when adjusting for outliers? A log transformation can be used.

```{r}
flights_2018$Carrier_Name <- as_factor(flights_2018$Carrier_Name)

flights_2018 %>%
  mutate(Total_Delay = Carrier_Delay + Late_Ac_Delay) %>%
  filter(Total_Delay>0) %>%
  mutate(Total_Delay_Log = log(Total_Delay)) %>%
  group_by(Carrier_Name) %>%
  summarise(Total_Delay_Mean_Log = mean(Total_Delay_Log)) %>%
  mutate(Carrier_Name = fct_reorder(Carrier_Name,Total_Delay_Mean_Log)) %>%
  ggplot(aes(x = Carrier_Name, y = Total_Delay_Mean_Log)) +
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#Comparing two airlines: Total Delay

```{r}
carrier2 <- "Southwest Airlines Co."
carrier1 <- "SkyWest Airlines Inc."

flights_2018 %>%
  mutate(Total_Delay = Carrier_Delay + Late_Ac_Delay) %>%
  filter(Total_Delay>0) %>%
  filter(Carrier_Name == carrier1 | Carrier_Name == carrier2) %>%
  ggplot(aes(x = Carrier_Name, y = Total_Delay)) +
  geom_boxplot(aes(group = Carrier_Name)) + scale_y_log10() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

flights_2018 %>%
  mutate(Total_Delay = Carrier_Delay + Late_Ac_Delay) %>%
  filter(Total_Delay>0) %>%
  filter(Carrier_Name == carrier1 | Carrier_Name == carrier2) %>%
  mutate(Total_Delay_Log = log(Total_Delay)) %>%
  t.test(Total_Delay_Log~Carrier_Name, data = .) %>%
  tidy() -> tout_total_delay_carrier

exp(tout_total_delay_carrier$estimate)
exp(c(low = tout_total_delay_carrier$conf.low, high = tout_total_delay_carrier$conf.high))

```
When looking at the total delay, between carrier delay and late aircraft, it seems like one can expect to be delayed about 1.23 times longer on SkyWest than on SouthWest airlines. 